Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-02-03 WSSSPE3/report/wg_main_best_practices.tex

Core questions that will need to be explored are in reliability
(reproducibility), usability, extensibility, knowledge management and continuity
(transitions between people). Answers to these will guide us on how a software
tool becomes part of the core workflow of well identified users (stakeholders)
relating to tool success and hence sustainability.
%\katznote{prev sentence is complex and awkward} % see fix
Ideas  that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment (e.g., new hardware,
new methodology);
\item Dimensions of sustainability: economic, technical, environmental and
obsolescence. 
%\katznote{not sure what the prev. comment goes with} %social.
\end{itemize}

@@

\subsubsection{Plans}
%\todo{short text here - not bullets}
The creation of a best-practices document needs a large and diverse community
involved. We have enlisted over ten contributors from the attendees at the
WSSSPE3 and those on the mailing list. The primary mechanism for developing this
document will be to examine and analyze the success of several well known
community scientific software and organizations supporting scientific software.
We will attempt then to abstract general principles and best practices. Some of
the tools identified for such analysis are the general purpose PETSc toolkit for
linear system solution, NWChem for computational chemistry and the CIG
(Computational Infrastructure for Geodynamics) organization dedicated to
supporting an ensemble of related tools for the geodynamics community. We also
established a timeline and a rough outline (see
Appendix~\ref{sec:appendix_best_practices}) for the report.

>>>>

Core questions that will need to be explored are in reproducibility, 
reliability,  usability, extensibility, knowledge management, as well as continuity
(transitions between people). Answers to these questions will guide us on how a software
tool becomes part of the core workflow of well identified users (stakeholders)
relating to tool success and hence sustainability.
%\katznote{prev sentence is complex and awkward} % see fix
Ideas  that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment (e.g., new hardware,
new methodology);
\item Dimensions of sustainability: economic, technical, environmental and
obsolescence. 
%\katznote{not sure what the prev. comment goes with} %social.
\end{itemize}

@@

\subsubsection{Plans}
%\todo{short text here - not bullets}
The creation of a document on best practices needs a large and diverse community
involved. We have enlisted over ten contributors from the attendees at the
WSSSPE3 and those on the mailing list. The primary mechanism for developing this
document will be to examine and analyze the success of several well known
community scientific software and organizations supporting scientific software.
We will attempt then to abstract general principles and best practices. Some of
the tools identified for such analysis are the general purpose PETSc toolkit for
linear system solution, NWChem for computational chemistry and the CIG
(Computational Infrastructure for Geodynamics) organization dedicated to
supporting an ensemble of related tools for the geodynamics community. We also
established a timeline and a rough outline (see
Appendix~\ref{sec:appendix_best_practices}) for the report.
